The not so widening achievement gap: an interntational comparison of the evolution of the SES achievement gap from 2000-2015

### 2017-09-09 paper/paper.Rnw
Fixed all tables/added some analysis - fixing graphs

table_coming <-
  summary_table %>%
  select(-perc) %>%
  gather(cats, vals, -(name:escs_dummy)) %>%
  unite(all_vals, escs_dummy, cats, sep = "_") %>%
  spread(all_vals, vals) %>%
  select(name, country, `0_n`, `0_Mean`, `0_s.e.`, `1_n`, `1_Mean`, `1_s.e.`) %>%
  mutate(`0_n` = as.character(`0_n`),
         `1_n` = as.character(`1_n`)) %>%
  xtable::xtable()

@@

print(table_coming,
      scalebox = '0.75',
      floating = FALSE,
      add.to.row = addtorow,
      include.colnames = FALSE,
      include.rownames = FALSE,
      hline.after = nrow(table_coming),
      caption.placement = "top",
      caption = "SES sample size and ISCED composition")
@

@@

On the other hand, we do have some countries which show a decrease in the SES achievement gap. Poland decreased by about \Sexpr{diff_data$math[["Poland"]]} and Denmark by \Sexpr{diff_data$math[["Denmark"]]}. However, the most notable cases are the United States and Germany. These two countries show a very high level of dispersion in the year 2000 with SES gaps of over 2 standard deviations. But in the 15-year time trend both countries reduced the gaps by \Sexpr{diff_data$math[["United States"]]} and \Sexpr{diff_data$math[["Germany"]]} respectively. Their distinctively big gaps in 2000 also show up in the work of \citet{micklewright}. This corroborates the findings of \citet{reardon_portilla}, which found a decreasing gap for kindergarteners. This decline is evident also for 15 year olds, suggesting it might be more of an institutional change rather than a specific grade-level policy. Do note that the trendline is for both mathematics and literacy. If it were only for mathematics, then the trendline would probably be even more extreme. Despite this, pooling both subjects gives even more strength to the analysis given that the results stand for both subjects together \footnote{Same graph is availabe for reading upon request, where we find pretty much the same results if not even more steep/downward slopes.}.

@@

Generally speaking, we see that most countries increased their achievement gap over time. France had an average increase of about 80\% since 2000 varying down to 40\%, whereas Germany had a similar figure but decreasing. Many of the countries we saw before that didn't have a very steep slope, such as Hungary or Australia, in fact had increases of about 40\% of their gap. In contrast, we see that the U.S and Poland had also quite significant decreases of about 40\%. The benefit of presenting these estimates this way is that we can actually asses the uncertainty of each calculation, and we do find that some of these estimates have wide variability. Despite this, most countries show a clear sign of either decreasing or increasing.

@@

% Add confidence intervals
\begin{figure}
\begin{center}
<<rate_increase/decrease, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
# You left off here. Plot the gap for 1 and 0 for all countries in the graph before.
complete_data_topbottom %>%
  filter(country %in% countries) %>%
  mutate(country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  select(wave, country, matches("*._math$")) %>%
  gather(ses, gap_size, matches("^\\d_mean_math$")) %>%
  separate(ses, c("ses", "delete"), sep = 2) %>%
  mutate(ses = gsub("_", "", ses)) %>%
  filter(complete.cases(.)) %>%
  # gather(lower_bound, lower, matches("\\d_lower")) %>%
  # gather(upper_bound, upper, matches("\\d_upper"))
  ggplot(aes(as.factor(wave), gap_size, group = ses, colour = ses, shape = ses)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_line(alpha = 0.4) +
  geom_line(stat = "smooth", method = "lm", aes(group = 1),
            formula = y ~ splines::ns(x, 1), linetype = "longdash",
            colour = "black") +
  scale_y_continuous(name = "Standardized test scores (mean 0)") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  scale_shape_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  ggtitle("Evolution of the achievement gap by top/bottom groups") +
  # geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1, alpha = 0.4) +
  # geom_hline(yintercept = 0, linetype = "longdash") +
  # coord_cartesian(ylim = c(0, 3)) +
  facet_wrap(~ country, ncol = 5) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Evolution of the achievement gap for separate SES groups}
\end{center}
\end{figure}

@@

complete_data_topbottom %>%
  bind_rows(complete_data_topmid) %>%
  bind_rows(complete_data_midbottom) %>%
  filter(country %in% c("United States", "Denmark", "France", "Germany"), type_test == "math") %>%
  select(wave, country, type_test, type, contains("math")) %>%
  mutate(type = factor(type,
                       levels = c("90th/10th SES gap", "90th/50th SES gap", "50th/10th SES gap"),
                       ordered = TRUE)) %>%
  gather(score, value, -(wave:type)) %>%
  separate(score, c("ses", "score"), sep = 2) %>%
  spread(score, value) %>%
  ggplot(aes(as.factor(wave), mean_math, group = ses, colour = ses, shape = ses)) +
  geom_line(alpha = 0.4) +
  geom_point(size = 2, alpha = 0.4) +
  coord_cartesian(ylim = c(-1, 2)) +
  scale_y_continuous(name = "Standardized test scores (mean 0)") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  scale_shape_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  ggtitle("Evolution of the top/bottom groups by different gaps") +
  facet_grid(country ~ type) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Evolution of the SES gaps for top/bottom groups}
\end{center}
\end{figure}

@@

% Add confidence intervals
\begin{figure}
\begin{center}
<<country_achievement_and_inequality, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
data_to_plot %>%
  mutate(Mean = scale(Mean),
         country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  gather(type, value, difference, Mean) %>%
  filter(country %in% countries) %>%
  ggplot(aes(as.character(wave), value, group = type, colour = type, shape = type)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_line(alpha = 0.4) +
  scale_y_continuous(name = "Test scores in SD") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  scale_shape_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  ggtitle("Evolution of the top/bottom groups by different gaps") +
  facet_wrap(~ country, ncol = 5) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Time trends of achievement gap and average achievement between 2000 and 2015}
\end{center}
\end{figure}

@@

<<complete_tracking_results, include = TRUE>>=
all_track_models %>%
  stan_table() %>%
  mutate(" " = c(
    "Only 1 track",
    "Age selection >= 15",
    "% of curric tracked",
    "Vocational Index",
    "Intercept"
  )) %>%
  xtable() %>%
  print(
  include.rownames = FALSE,
  add.to.row = table_details(all_track_models)
  )
@

@@

new_data %>%
  left_join(sum_ready_data) %>%
  mutate(age_selection = ifelse(age_selection == 1, ">= 15", "< 15")) %>%
  ggplot(aes(num_tracks, avg_diff, fill = num_tracks)) +
  geom_point(colour = "red") +
  geom_errorbar(aes(ymin = diff_low, ymax = diff_high), width = 0.2, colour = "red") +
  geom_col(aes(y = pred, alpha = 0.4)) +
  geom_errorbar(aes(ymin = pred_low, ymax = pred_high), width = 0.2) +
  facet_wrap(country ~ age_selection) +
  scale_y_continuous(name = "90/10 SD gap", limits = c(0, 4)) +
  scale_alpha(guide = F) +
  scale_x_discrete(name = "Number of tracks", labels = c("> 1 track", "1 track")) +
  scale_fill_manual(values = c("orangered2", "blue3"), guide = F) +
  ggtitle("Predicted 90/10 gap for different combinations of number of tracks and age of selection")
@

The red point for Germany is in the '\textgreater 1 track' and '\textless 15 age' setup, meaning that Germany has a tracking system that has more than 1 track and an age of selection of below 15. We see that the model predicts the level of inequality for that set up quite accurately, relative to the actual level (red point). But if the setup would be in the ideal '1 track' and age of selection above or equal to 15, the inequality would be much lower.

These graph reveals some interesting patterns. Generally speaking, we find that the simulation shows that virtually all countries which are not in this 'ideal' tracking, that would switch to the 'ideal' setup would see a reduction of the achievement gap. Yet we also find the opposite result. Countries which are in the 'ideal' tracking setup that would change to the 'worst' setup, all countries would see a widening of the achievement gap. On average, we find that if all countries switched to the 'ideal' setup, we would see an average reduction of 20\%. This is a very important figure considering that \citet{reardon2011} found that the US gap increased by about 40-50\% in 30 years. Had all countries in the 'ideal' tracking switched to the 'worst' tracking, we would see an average widening of nearly 72\%. For those countries which are not in the 'ideal' tracking setup, we compute below the percentage change from switching to the ideal tracking.

@@

final_avg_diff_table <-
  avg_diff_data %>%
  left_join(pred_avg_diff %>%
              filter(country %in% track_should_change & (num_tracks == 1 & age_selection == 1)) %>%
              select(country, pred), by = "country") %>%
  select(country, combn, avg_diff, pred) %>%
  separate(combn, c("num_tracks", "age_selection"), sep = "_") %>%
  transmute(
    Country = country,
    `Number of tracks` = ifelse(num_tracks == 0, "More than one track", "One track"),
    `Age of selection` = ifelse(age_selection == 1, "15 or more", "Less than 15"),
    `Inequality` = avg_diff %>% round(3),
    `Predicted` = pred %>% round(3),
    `% reduction` = (((`Inequality` - `Predicted`)/ `Inequality`) * 100) %>%
      ceiling %>% paste0("%")
    )

xtable(final_avg_diff_table) %>%
  print(
  include.rownames = FALSE
  )

@@

@@

# This is for when calculating the mean 'widening' of the gap if countries with 'ideal' tracking swtiched to the 'worse' tracking

@@

<<interaction_model, results = 'hide'>>=

@@

ready_data <- dif_data(complete_gaps, tracking_data)

<<interaction_table,  include = TRUE>>=
stan_models %>%
  stan_table() %>%
  mutate(" " = c(
    "Tracking Index",
    "Vocational Index",
    "Trackin * Vocational Index",
    "Intercept"
  )) %>%
  xtable() %>%
  print.xtable(
    include.rownames = FALSE,
    add.to.row = table_details(stan_models)
  )
@

@@

plot(plot_interaction, plot = FALSE)[[1]] +
  labs(x = "Tracking Index (SD)", y = "90/10 Achievement gap (SD)") +
  scale_colour_discrete(
    name = "Vocational index \n quantiles",
    labels = seq(100, 0, -25) %>% paste0("%")
    ) +
  scale_fill_discrete(guide = FALSE) +
  ggtitle("Interaction between Tracking and Vocational index")

@@

% Talk about the mechanisms why tracking might be explaining this, although tracking doesn't change over time.

% Because it might've caused it at the beginning and now it just exarcebates it.

% Because when combined with vocational enrollment, it might be exarcebarting even more, if enrollment is not high.

@@

<<cumulative_modeling_table, include = TRUE>>=
all_gaps_models_cum %>%
  stan_table() %>%
  mutate(" " = c(
    "Track Index",
    "Vocational Index",
    "Intercept"
  )) %>%
  setNames(c(" ", gsub(" SES gap", "", gaps))) %>%
  xtable() %>%
  print.xtable(
    include.rownames = FALSE,
    add.to.row = table_details(all_gaps_models_cum)
  )

>>>>

table_coming <-
  summary_table %>%
  select(-perc) %>%
  gather(cats, vals, -(name:escs_dummy)) %>%
  unite(all_vals, escs_dummy, cats, sep = "_") %>%
  spread(all_vals, vals) %>%
  select(name, country, `0_n`, `0_Mean`, `0_s.e.`, `1_n`, `1_Mean`, `1_s.e.`) %>%
  mutate(`0_n` = as.character(`0_n`),
         `1_n` = as.character(`1_n`))

@@

xtable(table_coming,
       caption = "SES sample size and ISCED composition"
       ) %>%
  print(
    add.to.row = addtorow,
    include.colnames = FALSE,
    include.rownames = FALSE,
    hline.after = nrow(table_coming)
  )
@

@@

On the other hand, we do have some countries which show a decrease in the SES achievement gap. Poland decreased by about \Sexpr{diff_data$math[["Poland"]]} and Denmark by \Sexpr{diff_data$math[["Denmark"]]}. However, the most notable cases are the United States and Germany. These two countries show a very high level of dispersion in the year 2000 with SES gaps of over 2 standard deviations. But in the 15-year time trend both countries reduced the gaps by \Sexpr{diff_data$math[["United States"]]} and \Sexpr{diff_data$math[["Germany"]]} respectively. Their distinctively big gaps in 2000 also show up in the work of \citet{micklewright}. This corroborates the findings of \citet{reardon_portilla}, which found a decreasing gap for kindergarteners. This decline is evident also for 15 year olds, suggesting it might be more of an institutional change rather than a specific grade-level policy. Do note that the trendline is for both mathematics and literacy. If it were only for mathematics, then the trendline would probably be even more extreme. Despite this, pooling both subjects gives even more strength to the analysis given that the results stand for both subjects together \footnote{Same graph is availabe for reading upon request, where we find pretty much the same results if not even more steep/downward slopes.}. We also plot the 80/20 and 70/30 SES gaps and find that very similar patterns as in the 90/10 SES gap (see figure 10 in the appendix).

@@

Generally speaking, we see that most countries increased their achievement gap over time. France had an average increase of about 80\% since 2000 varying down to 40\%, whereas Germany had a similar figure but decreasing. Many of the countries we saw before that didn't have a very steep slope, such as Hungary or Australia, in fact had increases of about 40\% of their gap. In contrast, we see that the U.S and Poland had also quite significant decreases of about 40\%. The benefit of presenting these estimates this way is that we can actually asses the uncertainty of each calculation, and we do find that some of these estimates have wide variability. Despite this, most countries show a clear sign of either decreasing or increasing. We plot the same graph for the 80/20 and 70/30 SES gaps in the appendix and the results hold for these achievement gaps as well.

@@

% Add confidence intervals
\begin{figure}
\begin{center}
<<rate_increase/decrease, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
# You left off here. Plot the gap for 1 and 0 for all countries in the graph before.
complete_data_topbottom %>%
  filter(country %in% countries) %>%
  mutate(country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  select(wave, country, matches("*._math$")) %>%
  gather(ses, gap_size, matches("^\\d_mean_math$")) %>%
  separate(ses, c("ses", "delete"), sep = 2) %>%
  mutate(ses = gsub("_", "", ses)) %>%
  filter(complete.cases(.)) %>%
  # gather(lower_bound, lower, matches("\\d_lower")) %>%
  # gather(upper_bound, upper, matches("\\d_upper"))
  ggplot(aes(as.factor(wave), gap_size, group = ses, colour = ses, shape = ses)) +
  geom_point(size = 2) +
  geom_line() +
  geom_line(stat = "smooth", method = "lm", aes(group = 1),
            formula = y ~ splines::ns(x, 1), linetype = "longdash",
            colour = "black") +
  scale_y_continuous(name = "Standardized test scores (mean 0)") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  scale_shape_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  ggtitle("Evolution of the achievement gap by top/bottom groups") +
  # geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1, alpha = 0.4) +
  # geom_hline(yintercept = 0, linetype = "longdash") +
  # coord_cartesian(ylim = c(0, 3)) +
  facet_wrap(~ country, ncol = 5) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Evolution of the achievement gap for separate SES groups}
\end{center}
\end{figure}

@@

complete_data_topbottom %>%
  bind_rows(complete_data_topmid) %>%
  bind_rows(complete_data_midbottom) %>%
  filter(country %in% c("United States", "Denmark", "France", "Germany"), type_test == "math") %>%
  select(wave, country, type_test, type, contains("math")) %>%
  mutate(type = factor(type,
                       levels = c("90th/10th SES gap", "90th/50th SES gap", "50th/10th SES gap"),
                       ordered = TRUE)) %>%
  gather(score, value, -(wave:type)) %>%
  separate(score, c("ses", "score"), sep = 2) %>%
  spread(score, value) %>%
  ggplot(aes(as.factor(wave), mean_math, group = ses, colour = ses, shape = ses)) +
  geom_line() +
  geom_point(size = 2) +
  coord_cartesian(ylim = c(-1, 2)) +
  scale_y_continuous(name = "Standardized test scores (mean 0)") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  scale_shape_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  ggtitle("Evolution of the top/bottom groups by different gaps") +
  facet_grid(country ~ type) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Evolution of the SES gaps for top/bottom groups}
\end{center}
\end{figure}

@@

% Add confidence intervals
\begin{figure}
\begin{center}
<<country_achievement_and_inequality, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
data_to_plot %>%
  mutate(Mean = scale(Mean),
         country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  gather(type, value, difference, Mean) %>%
  filter(country %in% countries) %>%
  ggplot(aes(as.character(wave), value, group = type, colour = type, shape = type)) +
  geom_point(size = 2) +
  geom_line() +
  scale_y_continuous(name = "Test scores in SD") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  scale_shape_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  ggtitle("Evolution of the relationship between achievement and inequality") +
  facet_wrap(~ country, ncol = 5) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Time trends of achievement gap and average achievement between 2000 and 2015}
\end{center}
\end{figure}

@@

<<complete_tracking_results, include = TRUE>>=
all_track_models %>%
  stan_table() %>%
  mutate(" " = c(
    "Only 1 track",
    "Age selection >= 15",
    "% of curric tracked",
    "Vocational Index",
    "Intercept"
  )) %>%
  xtable(
    caption = "Explaining 90/10 achievement gap - all countries pooled - Multilevel model"
  ) %>%
  print(
  include.rownames = FALSE,
  add.to.row = table_details(all_track_models)
  )
@

@@

new_data %>%
  left_join(sum_ready_data) %>%
  mutate(age_selection = ifelse(age_selection == 1, ">= 15", "< 15")) %>%
  ggplot(aes(num_tracks, avg_diff, fill = num_tracks)) +
  geom_col(aes(y = pred, colour = num_tracks)) +
  geom_errorbar(aes(ymin = pred_low, ymax = pred_high), width = 0.2) +
  geom_point(colour = "red", shape = 8) +
  geom_errorbar(aes(ymin = diff_low, ymax = diff_high), width = 0.2, colour = "red") +
  facet_wrap(country ~ age_selection) +
  scale_y_continuous(name = "90/10 SD gap", limits = c(0, 4)) +
  scale_alpha(guide = F) +
  scale_x_discrete(name = "Number of tracks", labels = c("> 1 track", "1 track")) +
  scale_fill_manual(values = c("white", "white"), guide = F) +
  scale_colour_manual(values = c("black", "black"), guide = F) +
  ggtitle("Predicted 90/10 gap for different combinations of number of tracks and age of selection")
@

The star point for Germany is in the '\textgreater 1 track' and '\textless 15 age' setup, meaning that Germany has a tracking system that has more than 1 track and an age of selection of below 15. We see that the model predicts the level of inequality for that set up quite accurately (height of the bar), relative to the actual level (the star point). But if the setup would be in the ideal '1 track' and age of selection above or equal to 15, the inequality would be much lower.

These graph reveals some interesting patterns. Generally speaking, we find that the simulation predicts that virtually all countries which are not in this 'ideal' tracking, that would switch to the 'ideal' setup, would experience a reduction of the achievement gap. Yet we also find the opposite result. Countries which are in the 'ideal' setup that would change to the 'worst' setup would see a widening of the achievement gap. On average, we find that if all countries switched to the 'ideal' setup, we would see an average reduction of 20\%. This is a very important figure considering that \citet{reardon2011} found that the US gap increased by about 40-50\% in 30 years. Had all countries in the 'ideal' tracking switched to the 'worst' tracking, we would see an average widening of nearly 72\%. For those countries which are not in the 'ideal' tracking setup, we compute below the percentage change from switching to the ideal tracking.

@@

final_avg_diff_table <-
  avg_diff_data %>%
  left_join(pred_avg_diff %>%
              filter(country %in% track_should_change & (num_tracks == 1 & age_selection == 1)) %>%
              select(country, pred), by = "country") %>%
  select(country, combn, avg_diff, pred) %>%
  separate(combn, c("num_tracks", "age_selection"), sep = "_") %>%
  transmute(
    Country = country,
    `Number of tracks` = ifelse(num_tracks == 0, "More than one track", "One track"),
    `Age of selection` = ifelse(age_selection == 1, "15 or more", "Less than 15"),
    `Inequality` = avg_diff %>% round(3),
    `Predicted` = pred %>% round(3),
    `% reduction` = (((`Inequality` - `Predicted`)/ `Inequality`) * 100) %>%
      ceiling %>% paste0("%")
    )
final_avg_diff_table %>%
  xtable(
    caption = "Simulation: reduction of achievement gap if countries switched to 'ideal' tracking"
  ) %>%
  print(
  include.rownames = FALSE
  )

@@

% This is for when calculating the mean 'widening' of the gap if countries with 'ideal' tracking swtiched to the 'worse' tracking

@@

@@

<<interaction_model, results = 'hide'>>=
ready_data <- dif_data(complete_gaps, tracking_data)

<<interaction_table,  include = TRUE>>=
stan_models %>%
  stan_table() %>%
  mutate(" " = c(
    "Tracking Index",
    "Vocational Index",
    "Trackin * Vocational Index",
    "Intercept"
  )) %>%
  xtable(
    caption = "What explains 90/10 achievement gap? Tracking and Vocational interaction"
  ) %>%
  print.xtable(
    include.rownames = FALSE,
    add.to.row = table_details(stan_models)
  )
@

<<coefficents_sif_diff, eval = F>>=

@@

sig(0.41, , -0.22, ) # the two arguments missing are the SE's of both coefficients.
# take from the model from above and finish calculate (already did before, they're sig different from each other but I don't want to rerun again! :P)

@@

data_interaction <-
  plot(plot_interaction, plot = FALSE)[[1]] %>% ggplot_build() %>% pluck(1) %>% pluck(1)

order_colour <-
  data_interaction %>%
  distinct(group, colour) %>%
  mutate(colour = rev(colour)) %>%
  pull(colour)

data_interaction %>%
  mutate(linetype = group %>% as.factor,
         colour = factor(colour, levels = order_colour)) %>%
  ggplot(aes(x, y)) +
  geom_line(aes(colour = colour, linetype = linetype), size = 1) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax, fill = colour), alpha = 0.4) +
  labs(x = "Tracking Index (SD)", y = "90/10 Achievement gap (SD)") +
  scale_colour_discrete(
    name = "Vocational index \n quantiles",
    labels = seq(100, 0, -25) %>% paste0("%")
  ) +
  scale_linetype_discrete(
    name = "Vocational index \n quantiles",
    labels = seq(100, 0, -25) %>% paste0("%")
  ) +
  scale_fill_discrete(guide = FALSE) +
  ggtitle("Interaction between Tracking and Vocational index")

@@

The exact mechanisms through whick the tracking setup is explaining these inequalities is still very speculative. We could argue that tracking might've introduced the inequality once it was implemented. Another possibility is that tracking, althought not changing, is currently exarcebating it every year. Another explanation, which we find strong evidence for, is that the tracking setup interacts very closely with the vocational setup. These explanations are very speculative and further research should carry the task of looking for a more concrete explanation.

@@

<<cumulative_modeling_table, include = TRUE>>=
all_gaps_models_cum %>%
  stan_table() %>%
  mutate(" " = c(
    "Track Index",
    "Vocational Index",
    "Intercept"
  )) %>%
  setNames(c(" ", gsub(" SES gap", "", gaps))) %>%
  xtable(
    caption = "What explains the cumulative achievement gaps? Models for three different achievement gaps"
  ) %>%
  print.xtable(
    include.rownames = FALSE,
    add.to.row = table_details(all_gaps_models_cum)
  )
